---
title: "Statistical inference with the GSS data"
output: 
  html_document: 
    fig_height: 4
    highlight: pygments
    theme: spacelab
---

## Setup

### Load packages

```{r load-packages, message = FALSE}
library(ggplot2)
library(dplyr)
library(statsr)
```

### Load data

```{r load-data}
load("gss.Rdata")

```



* * *

## Part 1: Data
From reading the documentation it states that the data was collected by using full probability sampling (random sampling) and block quota sampling. The mode of data collection were computer-assisted personal interview (CAPI), face-to-face interviews, telephone interviews.

The use of full probability sampling will achieve a casual inference, generalized to the whole population as all elements in the target population will have an equal chance of being selected for inclusion in the study. (random sampling)

The use of block quota sampling will infer a causal inference, only for the sample and will not generalize to the population due to this method not being random sampling.

* * *

## Part 2: Research question

in this part i want to see if there is a relation between gender (sex) and years of education (educ).

* * *

## Part 3: Exploratory data analysis
lets make a data frame which containes sex and education 
```{r}
gss_SE <- gss %>%
  select(sex, educ)
```
now it's better to get a summary from our data
```{r}
summary(gss_SE)
```
Now I want to see frequency of each gender
```{r}
gss_SE %>%
  group_by(sex) %>% 
  summarise(count = n()) %>%
  mutate(freq =count/ sum(count)*100)
```
And the same for education
```{r}
gss_SE %>%
  group_by(educ) %>%
  summarise(count = n()) %>%
  mutate(freq2 = count / sum(count) * 100)
```
As we see above, 12 years of study has the most frequency and it is kind of obvious as 12 years of education means finishing highschool that most of people do it. 


```{r}
gss_SE %>%
filter(!is.na(educ), sex == "Male" ) %>%
summarise(miuM = mean(educ), sdM = sd(educ), minM = min(educ), maxM = max(educ))

gss_SE %>%
filter(!is.na(educ), sex == "Female" ) %>%
summarise(miuF = mean(educ), sdF = sd(educ), minF = min(educ), maxF = max(educ))
```
From now on, I ploted some charts to help me have better vision of study and interprate accurately.
As we see below, both educations have left skewed distribution which means we have more data in left side of distribution. Male's education has bigger varible. 
```{r}
ggplot(gss_SE, aes(sex, educ)) + geom_boxplot()
```
```{r}
ggplot(gss_SE, aes(x = sex)) + geom_bar() 
```

As we can see below, 12 years of education has the most number among the others. 
```{r}
ggplot(gss_SE, aes(x = educ)) + geom_bar() 

```

```{r}
gss_SE %>%
filter(sex == "Male") %>%
ggplot(aes(x = educ)) + geom_bar()
```
```{r}
gss_SE %>%
  filter(sex == "Female", !is.na(educ)) %>%
  ggplot(aes(x= educ)) + geom_bar()
```
```{r}
gss_SE %>%
  filter(!is.na(educ)) %>%
  ggplot() +
    geom_bar(aes(x = factor(sex), fill = factor(educ)), position = "fill")
```

* * *

## Part 4: Inference

In this section, I want to test my hypothesis
Several conditions should be considered:
State hypotheses
H0: mu_Male = mu_Female

HA: mu_Male != mu_Female

The sampling method for each sample is simple random sampling.

The samples are independent.

Each population is at least 20 times larger than its respective sample.

The sampling distribution is approximately normal, which is generally the case if any of the following 
The sample size is greater than 40, without outlines.


State the method(s) to be used and why and how
Since we are comparing two sample means we will first calculate a confidence interval of two means and then calculate a two-sample t-test hypothesis test to test the above hypothesis stated above.

The confidence Interval will we allow us to see the difference in mean total years of study between men and women. If the confidence interval is between the null value then we can determine that there is no statistically meaningful difference in the means.

The two-sample t-test to determine whether the difference between means found in the sample is significantly different from the hypothesized difference between means.

A Significance level of 0.05 will be applied.

Perform inference
Cofidence Interval


```{r}
gss_SE %>%
inference(y = educ , x = sex, type = "ci", statistic = "mean",
          method = "theoretical" , conf_level = 0.95)
```


Hypothesis Test

```{r}
gss_SE %>%
inference(y = educ , x = sex, type = "ht", statistic = "mean",
          method = "theoretical", alternative = "twosided" , null = 0, sig_level = 0.05)
```


Interpret results
Cofidence Interval
With 95% confidence the difference in mean total years of study between men and women is between 0.2001 and 0.3067 units. We estimate that the difference in mean total years of study is between 0.2001 and 0.3067 units with men having the higher values. There is a small, but statistically meaningful difference in the means. Since this 95% confidence interval does not include a null, we can conclude that there is a statistically significant difference in mean total years of study between men and women.

Hypothesis Test
A P value of < 0.0001 was observed, with a 0.05 significance level we can reject the null hypothesis and conclude there is a statistically meaningful difference in mean total years of study between men and women.

Due to the sample size being quite large, the p value was shrunk. The effect size observed is quite small but I still feel that the statistically significant finding is meaningful due to the large sample being closer to the population.





